---
title: "R_LEGO_PROJECT"
author: "Justyna Frączek"
date: "2023-12-04"
output:
  html_document: 
    toc: yes
    toc_float: yes
    theme: readable
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggsci)
library(shiny)
library(plotly)
library(flexdashboard)
library(tidyverse)
library(readr)
```

```{r load_data, echo=FALSE}

load_csv_gz_data <- function(file_path) {
    data <- read.csv(file_path)
  return(data)
}

colors <- load_csv_gz_data("rebrickable/colors.csv.gz")
elements <- load_csv_gz_data("rebrickable/elements.csv.gz")
inventories <- load_csv_gz_data("rebrickable/inventories.csv.gz")
inventory_minifigs <- load_csv_gz_data("rebrickable/inventory_minifigs.csv.gz")
inventory_parts <- load_csv_gz_data("rebrickable/inventory_parts.csv.gz")
inventory_sets <- load_csv_gz_data("rebrickable/inventory_sets.csv.gz")
minifigs <- load_csv_gz_data("rebrickable/minifigs.csv.gz")
part_categories <-load_csv_gz_data("rebrickable/part_categories.csv.gz")
part_relationships <-load_csv_gz_data("rebrickable/part_relationships.csv.gz")
parts <-load_csv_gz_data("rebrickable/parts.csv.gz")
sets <-load_csv_gz_data("rebrickable/sets.csv.gz")
themes <- load_csv_gz_data("rebrickable/themes.csv.gz")

```

```{r remove_missing_data, echo=FALSE}
remove_null <- function(data) {
  data_without_missing_vals <- na.omit(data)
  return(data_without_missing_vals)
}

```
## R Executive summary

```{r summary, echo=FALSE}

#head(colors)
#head(elements)
#head(inventories)
#head(inventory_minifigs)
#head(inventory_parts)
#head(inventory_sets)
#head(minifigs)
#head(part_categories)
#head(part_relationships)
#head(parts)
#head(sets)
#head(themes)

colors_no_null <- remove_null(colors)
elements_no_null <- remove_null(elements)
inventories_no_null <- remove_null(inventories)
inventory_minifigs_no_null <- remove_null(inventory_minifigs)
inventory_parts_no_null <- remove_null(inventory_parts)
inventory_sets_no_null <- remove_null(inventory_sets)
minifigs_no_null <- remove_null(minifigs)
part_categories_no_null <- remove_null(part_categories)
part_relationships_no_null <- remove_null(part_relationships)
parts_no_null <- remove_null(parts)
sets_no_null <- remove_null(sets)
themes_no_null <- remove_null(themes)

head(colors_no_null)
head(elements_no_null)
head(inventories_no_null)
head(inventory_minifigs_no_null)
head(inventory_parts_no_null)
head(inventory_sets_no_null)
head(minifigs_no_null)
head(part_categories_no_null)
head(part_relationships_no_null)
head(parts_no_null)
head(sets_no_null)
head(themes_no_null)

```

## Basic statistics

### Sizes and basic statistics of data sets after removing null values: 

#### Colors dataset:

``` {r colors, echo=FALSE}
dim(colors_no_null)
summary(colors_no_null)
```

#### Elements dataset:

``` {r elements, echo=FALSE}
dim(elements_no_null)
summary(elements_no_null)
```

#### Inventories dataset:

``` {r inventories, echo=FALSE}
dim(inventories_no_null)
summary(inventories_no_null)
```

#### Inventory minifigs dataset:

``` {r inventory_minifigs, echo=FALSE}
dim(inventory_minifigs_no_null)
summary(inventory_minifigs_no_null)
```

#### Inventory parts dataset:

``` {r inventory_parts, echo=FALSE}
dim(inventory_parts_no_null)
summary(inventory_parts_no_null)
```

#### Inventory sets dataset:

``` {r inventory_sets, echo=FALSE}
dim(inventory_sets_no_null)
summary(inventory_sets_no_null)
```

#### Minifigs dataset:

``` {r minifigs, echo=FALSE}
dim(minifigs_no_null)
summary(minifigs_no_null)
```

#### Part categories dataset:

``` {r part_categories, echo=FALSE}
dim(part_categories_no_null)
summary(part_categories_no_null)
```

#### Part relationships dataset:

``` {r part_relationships, echo=FALSE}
dim(part_relationships_no_null)
summary(part_relationships_no_null)
```

#### Parts dataset:

``` {r parts, echo=FALSE}
dim(parts_no_null)
summary(parts_no_null)
```

#### Sets dataset:

``` {r sets, echo=FALSE}
dim(sets_no_null)
summary(sets_no_null)
```

#### Themes dataset:

``` {r themes, echo=FALSE}
dim(themes_no_null)
summary(themes_no_null)
```

## Attributes analysis

### Top 10 years with most sets released based on inventory sets quantity

``` {r top_sets, echo=FALSE}
inner_merged_data <- merge(sets_no_null, inventory_sets_no_null, by = "set_num")

yearly_total_quantity <- inner_merged_data %>% 
  group_by(year) %>% 
  summarise(TotalQuantity = sum(quantity))
  
top_10_years <- yearly_total_quantity %>%
  top_n(10, TotalQuantity) %>%
  arrange(desc(TotalQuantity))


ggplot(top_10_years, aes(x = as.factor(year), y = TotalQuantity, fill = year)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  theme_minimal() +
  labs(x = "Year", y = "Number of Sets Released") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  


```

### Most common theme for each of these 10 years based on inventory sets quantity

``` {r theme_quantity, echo=FALSE}

merged_themes <- merge(sets_no_null, themes_no_null, by.x = "theme_id", by.y = "id")
merged_data <- merge(merged_themes, inventory_sets_no_null, by = "set_num")
#head(merged_data)

#move to top
top_10_years_with_the_most_sets <- inner_merged_data %>%
  group_by(year) %>% 
  summarise(TotalQuantity = sum(quantity)) %>%
  top_n(10, TotalQuantity) %>%
  arrange(desc(TotalQuantity)) %>%
  pull(year)


years_with_most_sets <- merged_data %>%
  filter(year %in% top_10_years_with_the_most_sets)

most_common_themes_in_each_year <- years_with_most_sets %>%
  group_by(year, name.y) %>%
  summarise(TotalQuantity = sum(quantity)) %>%
  group_by(year) %>%
  filter(TotalQuantity == max(TotalQuantity)) 

ggplot(most_common_themes_in_each_year, aes(x = factor(year), y = TotalQuantity, fill = name.y)) +
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Number of sets' themes in each year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Variables corellations

``` {r inventory_parts_correlation, echo=FALSE}

```
## Summary the most important trends

## Prediction of future Lego trends