---
title: "R_LEGO_PROJECT"
author: "Justyna Frączek"
date: "2023-12-04"
output:
  html_document: 
    toc: yes
    toc_float: yes
    theme: readable
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggsci)
library(shiny)
library(plotly)
library(flexdashboard)
library(tidyverse)
library(readr)
library(corrplot)
library(dplyr)
library(tidyr)
library(prophet)
```

```{r load_data, echo=FALSE, cache=TRUE}

load_csv_gz_data <- function(file_path) {
    data <- read.csv(file_path)
  return(data)
}

colors <- load_csv_gz_data("rebrickable/colors.csv.gz")
elements <- load_csv_gz_data("rebrickable/elements.csv.gz")
inventories <- load_csv_gz_data("rebrickable/inventories.csv.gz")
inventory_minifigs <- load_csv_gz_data("rebrickable/inventory_minifigs.csv.gz")
inventory_parts <- load_csv_gz_data("rebrickable/inventory_parts.csv.gz")
inventory_sets <- load_csv_gz_data("rebrickable/inventory_sets.csv.gz")
minifigs <- load_csv_gz_data("rebrickable/minifigs.csv.gz")
part_categories <-load_csv_gz_data("rebrickable/part_categories.csv.gz")
part_relationships <-load_csv_gz_data("rebrickable/part_relationships.csv.gz")
parts <-load_csv_gz_data("rebrickable/parts.csv.gz")
sets <-load_csv_gz_data("rebrickable/sets.csv.gz")
themes <- load_csv_gz_data("rebrickable/themes.csv.gz")

```

```{r remove_missing_data, echo=FALSE, cache=TRUE}
remove_null <- function(data) {
  data_without_missing_vals <- na.omit(data)
  return(data_without_missing_vals)
}

```


```{r yearly_sets_total_quantity, echo=FALSE, cache=TRUE}
yearly_sets_total_quantity <- function(data) {
  yearly_total_quantity <- data %>% 
  group_by(year) %>% 
  summarise(TotalQuantity = sum(quantity))
}

```
## R Executive summary

This comprehensive analysis examines the temporal evolution of LEGO set releases, exploring various dimensions such as set count, themes popularity, color variety, and complexity. The analysis utilizes data spanning several decades, emphasizing trends and fluctuations in the LEGO product strategy.

The LEGO Group has significantly increased the number of set releases per year since 2000, with notable market responsiveness observed during the COVID-19 pandemic, as seen in the peaks of 2020 and 2021.

The analysis identifies a surge in product diversity, highlighted by the growing number of unique colors per set, indicating a strategic pivot towards more visually engaging and complex offerings.

Thematic trends reveal that certain series, such as DC Super Heroes and various Minifigures series, dominate release counts, reflecting both sustained popularity and strategic brand partnerships.

An examination of the most popular colors shows a dominance of traditional colors, with black and white leading, suggesting these colors' foundational roles in set design.

The correlation between set size and color variety is positive, indicating that larger sets tend to offer a greater spectrum of colors, adding to their complexity and appeal.

## Basic statistics

```{r remove_null_vals, echo=FALSE, cache=TRUE}
colors_no_null <- remove_null(colors)
elements_no_null <- remove_null(elements)
inventories_no_null <- remove_null(inventories)
inventory_minifigs_no_null <- remove_null(inventory_minifigs)
inventory_parts_no_null <- remove_null(inventory_parts)
inventory_sets_no_null <- remove_null(inventory_sets)
minifigs_no_null <- remove_null(minifigs)
part_categories_no_null <- remove_null(part_categories)
part_relationships_no_null <- remove_null(part_relationships)
parts_no_null <- remove_null(parts)
sets_no_null <- remove_null(sets)
themes_no_null <- remove_null(themes)

```

### Sizes and basic statistics of data sets after removing null values: 

#### Colors dataset:

Dataset size: 

``` {r colors, echo=FALSE, cache=TRUE}
dim(colors_no_null)
summary(colors_no_null)
```

#### Elements dataset:

Dataset size: 

``` {r elements, echo=FALSE, cache=TRUE}
dim(elements_no_null)
summary(elements_no_null)
```

#### Inventories dataset:

Dataset size: 

``` {r inventories, echo=FALSE, cache=TRUE}
dim(inventories_no_null)
summary(inventories_no_null)
```

#### Inventory minifigs dataset:

Dataset size: 

``` {r inventory_minifigs, echo=FALSE, cache=TRUE}
dim(inventory_minifigs_no_null)
summary(inventory_minifigs_no_null)
```

#### Inventory parts dataset:

Dataset size: 

``` {r inventory_parts, echo=FALSE, cache=TRUE}
dim(inventory_parts_no_null)
summary(inventory_parts_no_null)
```

#### Inventory sets dataset:

Dataset size: 
``` {r inventory_sets, echo=FALSE, cache=TRUE}
dim(inventory_sets_no_null)
summary(inventory_sets_no_null)
```

#### Minifigs dataset:

Dataset size: 

``` {r minifigs, echo=FALSE, cache=TRUE}
dim(minifigs_no_null)
summary(minifigs_no_null)
```

#### Part categories dataset:

Dataset size: 

``` {r part_categories, echo=FALSE, cache=TRUE}
dim(part_categories_no_null)
summary(part_categories_no_null)
```

#### Part relationships dataset:

Dataset size: 

``` {r part_relationships, echo=FALSE, cache=TRUE}
dim(part_relationships_no_null)
summary(part_relationships_no_null)
```

#### Parts dataset:

Dataset size: 

``` {r parts, echo=FALSE, cache=TRUE}
dim(parts_no_null)
summary(parts_no_null)
```

#### Sets dataset:

Dataset size: 

``` {r sets, echo=FALSE, cache=TRUE}
dim(sets_no_null)
summary(sets_no_null)
```

#### Themes dataset:

Dataset size: 

``` {r themes, echo=FALSE, cache=TRUE}
dim(themes_no_null)
summary(themes_no_null)
```

## Attributes analysis


### Number of sets released over years based on inventory sets quantity

```{r sets_slider, echo=FALSE, cache=TRUE}
inner_merged_data <- merge(sets_no_null, inventory_sets_no_null, by = "set_num")

yearly_total_quantity <- yearly_sets_total_quantity(inner_merged_data)

p <- plot_ly(yearly_total_quantity, x = ~year, y = ~TotalQuantity, type = 'scatter', mode = 'lines+markers') %>%
  layout(
    title = "Number of Sets Released Over Years",
    xaxis = list(
      title = "Year",
      rangeselector = list(
        buttons = list(
          list(count = 5, label = "5y", step = "year", stepmode = "backward"),
          list(count = 10, label = "10y", step = "year", stepmode = "backward"),
          list(count = 20, label = "20y", step = "year", stepmode = "backward"),
          list(step = "all", label = "All")
        )
      ),
      rangeslider = list(type = "date")
    ),
    yaxis = list(title = "Number of Sets")
  )

p

```

Conclusions:

We can observe a rapid increase in the number of sets released after 2000. The graph shows that the release frequency became more volatile, with significant peaks and troughs.

In the most recent years displayed on the graph, there is a notable fluctuation with sharp increases in the number of sets followed by declines. This could be due to various factors such as market strategy, changes driven by COVID-19 pandemic, where we can see a decent increase of sets in 2020 and 2021 years.

### Top 10 years with most sets released based on inventory sets quantity

``` {r top_sets, echo=FALSE, cache=TRUE}

  
top_10_years <- yearly_total_quantity %>%
  top_n(10, TotalQuantity) %>%
  arrange(desc(TotalQuantity))

ggplot(top_10_years, aes(x = as.factor(year), y = TotalQuantity, fill = year)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  theme_minimal() +
  labs(x = "Year", y = "Number of Sets Released") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  


```

### Most popular themes based on the number of sets released during these 10 years

``` {r theme_quantity, echo=FALSE, cache=TRUE}

merged_themes <- merge(sets_no_null, themes_no_null, by.x = "theme_id", by.y = "id")
merged_data <- merge(merged_themes, inventory_sets_no_null, by = "set_num")
#head(merged_data)


top_10_years_with_the_most_sets <- yearly_total_quantity %>%
  top_n(10, TotalQuantity) %>%
  arrange(desc(TotalQuantity)) %>%
  pull(year)


years_with_most_sets <- merged_data %>%
  filter(year %in% top_10_years_with_the_most_sets)

most_common_themes_in_each_year <- years_with_most_sets %>%
  group_by(year, name.y) %>%
  summarise(TotalQuantity = sum(quantity)) %>%
  group_by(year) %>%
  filter(TotalQuantity == max(TotalQuantity)) 

get_themes <- most_common_themes_in_each_year %>% pull(name.y)

ggplot(most_common_themes_in_each_year, aes(x = factor(year), y = TotalQuantity, fill = name.y)) +
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Number of sets with theme") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Most popular colors of elements for these themes

``` {r themes_color_quantity, echo=FALSE, cache=TRUE}

color_usage <- inventory_parts_no_null %>%
 inner_join(colors_no_null, c("color_id" = "id")) %>%
 inner_join(inventories_no_null, c("inventory_id" = "id")) %>%
 inner_join(sets_no_null, by = "set_num") %>%
 inner_join(themes_no_null, c("theme_id" = "id"))

themes_with_most_sets <- color_usage %>%
  filter(name %in% get_themes)

color_theme_usage <- themes_with_most_sets %>%
  group_by(name) %>%
  count(name.x) %>%
  top_n(1) %>%  
  ungroup() %>%
  select(name, n, name.x) %>%
  arrange(desc(n)) 

ggplot(color_theme_usage, aes(x = factor(name), y = n, fill = name.x)) +
  geom_bar(stat = "identity") +
  labs(x = "Theme name", y = "Number of elements in color") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

### 10 most popular colors of elements in general

``` {r color_quantity, echo=FALSE, cache=TRUE}

color_usage <- inventory_parts_no_null %>%
 inner_join(colors_no_null, c("color_id" = "id"))

color_count <- color_usage %>%
  group_by(name, rgb) %>%
  summarize(num_parts = n()) %>%
  arrange(desc(num_parts)) %>%
  head(n = 10)

ggplot(color_count, aes(x = factor(name), y = num_parts, fill = name)) +
 geom_bar(stat = "identity") +
 labs(x = "Color name", y = "Number of elements in color") + 
  theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
  
```

### Color variety in sets over the years

``` {r color_variety, echo=FALSE, cache=TRUE}

set_colors_data <- sets_no_null %>%
  inner_join(inventories_no_null, by="set_num") %>%
  inner_join(inventory_parts_no_null, by = c("id" = "inventory_id")) %>%
  inner_join(colors_no_null, by = c("color_id" = "id"))

set_colors_data$year <- as.numeric(as.character(set_colors_data$year))

color_variety_per_year <- set_colors_data %>%
  group_by(year, set_num) %>%
  summarize(unique_colors = n_distinct(color_id), .groups = 'drop') %>%
  group_by(year) %>%
  summarize(average_unique_colors = mean(unique_colors, na.rm = TRUE), .groups = 'drop')

ggplot(color_variety_per_year, aes(x = year, y = average_unique_colors)) +
  geom_line(color = "coral", linewidth = 1) +
  geom_point(color = "darkred") +
  labs(title = "Average Unique Colors per Lego Set Over Time",
       x = "Year",
       y = "Average Unique Colors") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

Conclusions:
  
The plot shows a clear upward trend in the average number of unique colors used in Lego sets from around the 1950s to the present. Notably, there is a significant increase starting in the early 2000s, where the average number of unique colors per set rises more steeply compared to previous decades.

This trend could be indicative of Lego's strategy to make sets more appealing and varied, perhaps in response to market demands for more intricate and visually stimulating products.

### Distribution of the number of minifigs in set - 10 sets with most minifigures

```{r minifigs_in_set2, echo=FALSE}
inventorysets_inventory_minifigs_data2 <- merge(inventory_sets_no_null, inventory_minifigs_no_null, by = "inventory_id")

inv_inv_mini = left_join(inventories_no_null, inventory_minifigs, by=c("id" = "inventory_id"))
i2_sets = left_join(inv_inv_mini, sets_no_null, by=c("set_num" = "set_num"))

minifigs_count_per_set2 <- i2_sets %>%
  group_by(id, set_num) %>%
  summarize(num_minifigs = n(),name = first(name), .groups = 'drop') %>%
  distinct(set_num, .keep_all = TRUE) %>%
  arrange(desc(num_minifigs)) %>%
  head(n = 10)

ggplot(minifigs_count_per_set2, aes(x = factor(paste(set_num,name)), y = num_minifigs, fill = name)) +
  geom_bar(stat = "identity") +
  labs(x = "Set name", y = "Number of minifigs") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.margin = margin(t = 30, r = 30, b = 30, l = 30, unit = "pt"))


minifigs_count_per_set2

```

### Number of minifigures included per set over time

```{r minifigs_in_set_over_time, echo=FALSE}

minifigs_per_set_year <- i2_sets %>%
  group_by(id, set_num) %>%
  summarize(num_minifigs = n(), year=first(year), .groups = 'drop') %>%
  distinct(set_num, .keep_all = TRUE) %>%
  ungroup()

minifigs_per_set_year <- minifigs_per_set_year %>%
  group_by(year) %>%
  summarize(average_minifigs = mean(num_minifigs, na.rm = TRUE), .groups = 'drop')

ggplot(minifigs_per_set_year, aes(x = year, y = average_minifigs)) +
  geom_line(color = "coral", linewidth = 1) +
  geom_point(color = "darkred") +
  labs(title = "Average number of minifigs per set over years",
       x = "Year",
       y = "Average number of minifigs per set") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

Conclusions:

For a significant period, specifically from the early years displayed up to around the late 1970s, the average number of minifigures per set remained relatively constant and close to 1.

Starting from the early 1980s, there is noticeable variability, with the average number of minifigures per set fluctuating more significantly. The fluctuations appear to be somewhat cyclical with peaks and troughs.

The trend becomes more pronounced in later years, with the variability increasing, which could be indicative of more diverse set offerings, special editions, or changes in set design philosophy.

## Variables corellations


### Correlation between size of a set and number of colors in a set
``` {r size_colors, echo=FALSE}

sets_colors <- sets_no_null %>%
  left_join(inventories_no_null, by = "set_num") %>%
  left_join(inventory_parts_no_null, by = c("id" = "inventory_id")) %>%
  group_by(set_num) %>%
  summarize(num_parts = sum(num_parts), num_colors = n_distinct(color_id)) %>%
  ungroup()

# Calculate the correlation matrix
cor_matrix <- cor(sets_colors[, c("num_parts", "num_colors")], use = "complete.obs")

# Plot the correlation matrix
corrplot(cor_matrix, method = "circle")

```

Conclusion:

The heatmap suggests that there is a positive correlation between the size of a Lego set (as measured by the number of parts) and the color diversity within the set (as measured by the number of unique colors). Sets that have a higher number of parts tend also to have a higher number of different colors.

### Correlation between the number of parts in a set and the complexity of a set

Complexity of a set can be achieved by approximate the number of unique part categories used in each set.

``` {r parts_complexity, echo=FALSE}

sets_parts <- sets_no_null %>%
  inner_join(inventories_no_null, by = "set_num") %>%
  inner_join(inventory_parts_no_null, by = c("id" = "inventory_id")) %>%
  inner_join(parts_no_null, by = "part_num")

category_complexity <- sets_parts %>%
  group_by(set_num) %>%
  summarize(num_unique_categories = n_distinct(part_cat_id), .groups = 'drop')

sets_complexity <- sets_no_null %>%
  inner_join(category_complexity, by = "set_num")

correlation_matrix <- cor(sets_complexity$num_parts, sets_complexity$num_unique_categories, use = "complete.obs")

ggplot(sets_complexity, aes(x = num_parts, y = num_unique_categories)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Correlation between Number of Parts and Complexity",
       x = "Number of Parts",
       y = "Number of Unique Categories")

print(correlation_matrix)

```

Conclusion:

The scatter plot reveals a positive correlation between the number of parts and set complexity. As the number of parts in a set increases, the number of unique part categories tends to increase as well, suggesting that larger sets are generally more complex. This relationship seems to hold strongly for sets with a smaller number of parts, as indicated by the dense cluster of points toward the origin, where the increase in complexity with the number of parts is quite pronounced.

For sets with a very high number of parts (toward the right end of the X-axis), the data points become more spread out, indicating more variability in complexity for these larger sets. It suggests that once a set reaches a certain size, the addition of more parts does not necessarily increase complexity at the same rate. This could be due to the use of repeated parts within these large sets or a design choice to not increase complexity despite a higher part count.


## Summary the most important trends

Over the years, the LEGO Group has navigated through various market trends and consumer preferences, as evidenced by the evolution of its set designs. The data shows an overall trend towards increased complexity and diversity in set offerings, which can be attributed to the following key developments:

### Innovation in Color and Complexity: 

There has been a clear trajectory towards incorporating a broader palette of colors in sets, coupled with an increase in the number of parts.

### Thematic Adaptability: 

LEGO has shown adaptability in its themes, often aligning with popular culture and consumer interests, as seen in the consistent popularity of certain minifigure series and licensed themes.

### Market Responsiveness: 

Fluctuations in set releases correlate with global events, such as the pandemic, where LEGO appeared to capitalize on increased demand for indoor activities, as well as adjust its product strategy in response to the economic climate.

### Consumer Engagement: 

The changing number of minifigures and the variability in sets suggest a strategic focus on consumer engagement, offering a mix of both collector-focused and play-intensive sets.


In conclusion, LEGO's ability to evolve while retaining its core principles has allowed it to remain a leader in the toy industry, consistently delivering products that resonate with consumers of all ages.

## Prediction of future Lego trends

I will use Prophet ML model to predict the number of released sets from the current year up to 2030.

``` {r future_prediction, echo=FALSE}

prophet_data <- yearly_total_quantity %>%
  mutate(ds = as.Date(paste(year, "-01-01", sep="")), y = TotalQuantity)

m <- prophet(prophet_data)

future <- make_future_dataframe(m, periods = 10, freq = "year")

forecast <- predict(m, future)

plot(m, forecast)

tail(forecast[c('ds', 'yhat', 'yhat_lower', 'yhat_upper')], 10)
```

Conclusion:

The prediction suggests a continued increase in the number of Lego sets released each year. 

If the trend observed in the past continues without significant change, we might expect to see a rise in the number of Lego sets released annually up to 2030. However, predictions should be taken with caution due to the potential impact of unforeseen future events.
